---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
set.seed(420)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 300,
  fig.path = "man/figures/",
  fig.width = 9,
  fig.height = 6,
  out.width = "70%"
)
```

# NLJ

The **nlj** package—named in honor of British statistician Norman Lloyd
Johnson—provides specialized tools for working with the unbounded Johnson SU
distribution. In addition to functions for density evaluation, distribution and
quantile computation, and random variate generation, the package offers
automatic normalization routines and a generalized transformation-based
regression model that leverages the inverse hyperbolic sine transformation.

---

## Installation

Install the development version of **nlj** from GitHub using **devtools**:

``` r
# install.packages("devtools")
devtools::install_github("shanedrabing/nlj")
```

---

## Overview of Functions and Examples

**nlj** provides a suite of tools for the Johnson SU distribution, automatic
data normalization, and advanced regression modeling. The following examples
demonstrate the package’s core functionality.

---

### Johnson-SU Distribution Functions

The Johnson-SU distribution functions—**djohnson**, **pjohnson**, **qjohnson**,
and **rjohnson**—provide density evaluation, cumulative distribution
computation, quantile inversion, and random sample generation. They support the
full Johnson-SU parameterization for flexible control of skewness and kurtosis.

#### Density

**djohnson** computes the Johnson-SU density for a specified set of input
values.

```{r djohnson}
x <- seq(-pi, pi, length.out = 300)
d <- nlj::djohnson(x, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
plot(x, d, type = "l",
     main = "Johnson-SU Density",
     xlab = "X", ylab = "Density")
```

#### Cumulative Distribution

**pjohnson** returns the cumulative distribution function (CDF) values for
specified quantiles under the Johnson-SU model.

```{r pjohnson}
q <- seq(-pi, pi, length.out = 300)
p <- nlj::pjohnson(q, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
plot(q, p, type = "l",
     main = "Johnson-SU Cumulative Distribution",
     xlab = "Quantile", ylab = "Probability")
```

#### Quantile Function

**qjohnson** obtains the quantile (inverse CDF) corresponding to specified
probability levels, facilitating probability-based analyses within the
Johnson-SU framework.

```{r qjohnson}
p <- seq(0, 1, length.out = 300)
q <- nlj::qjohnson(p, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
plot(p, q, type = "l",
     main = "Johnson-SU Quantile",
     xlab = "Probability", ylab = "Quantile")
```

#### Random Deviate Generation

**rjohnson** generates random samples from the Johnson-SU distribution, useful
for simulation, bootstrapping, or Monte Carlo studies.

```{r rjohnson}
nlj::rjohnson(5, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
```

---

### Automatic Normalization Functions

The package provides **znorm** and **zjohnson**, automated normalization
functions that compute and apply data transformations, track transformation
parameters, and support denormalization. These utilities streamline the process
of preparing data in standardized or distribution-adapted form.

#### Johnson-SU Normalization

**zjohnson** performs Johnson-SU normalization by estimating distribution
parameters that align the input data to a standard normal reference. It adapts
to skewness and heavy tails, producing a transformation that accommodates
non-Gaussian data characteristics.

```{r johnson-density}
# Example data: Wind speed
x <- sort(airquality$Wind)
d <- density(x)

# Apply Johnson-SU normalization
zj <- nlj::zjohnson(x)

# Extract the density estimates
xd <- d$x
yd <- d$y
yj <- zj$fd(xd)

# Plot the density comparison
plot(range(xd), range(c(yd, yj)), type = "n",
     main = "Wind Speed Density Estimation",
     xlab = "Wind Speed", ylab = "Density")
lines(xd, yd, col = "black")  # Original density
lines(xd, yj, col = "red")   # Johnson-SU normalized density
legend(min(xd), max(c(yd, yj)),
       c("Kernel", "Johnson-SU"),
       col = c("black", "red"),
       lwd = 1, bg = "white")
```

---

### Generalized Asinh Transformation Model (GATM)

**lm.gat** implements the Generalized Asinh Transformation Model, fitting
regression models with a parameterized inverse hyperbolic sine transform on
predictors. This approach can enhance model fit in the presence of
nonlinearity, particularly when linear models are insufficient.

#### Automatic Relationship Optimization

The following example illustrates **lm.gat** in action. We predict miles per
gallon (**mpg**) from engine horsepower (**hp**) using both a standard linear
model and a GATM model. The comparison highlights how the GATM transformation
captures non-linear patterns that a simple linear fit does not.

```{r gatm}
# Example data: Horsepower and miles per gallon (MPG)
i <- order(mtcars$hp)
x <- mtcars$hp[i]
y <- mtcars$mpg[i]

# Fit simple regression
m <- lm(y ~ x)

# Fit non-linear model
gat <- nlj::lm.gat(y ~ x, iterations = 3)

# Plot
plot(x, y, pch = 16,
     main = "Horsepower (HP) vs Miles Per Gallon (MPG)",
     xlab = "HP", ylab = "MPG")
lines(x, m$fitted.values, col = "red")
lines(x, gat$z, col = "blue")
legend(min(x), max(y),
       c("Simple", "GATM"),
       col = c("red", "blue"),
       lwd = 1, bg = "white")
```

#### Complex Nonlinear Interaction Optimization

The next example extends **lm.gat** to a model with multiple interaction terms,
demonstrating its ability to manage high-dimensional, non-linear relationships.
We illustrate this by modeling quarter-mile time (**qsec**) as a function of
**hp**, **mpg**, and **wt**, including all two-way and three-way interactions.

First, we fit a standard linear model with these terms and examine its summary.
Then, we apply **lm.gat** to the same formula, showcasing how the parameterized
asinh transformation can yield a superior fit by accommodating complex variable
interactions.

```{r interaction-simple}
# Fit simple regression
m <- lm(qsec ~ hp * mpg * wt, mtcars)

summary(m)
```

```{r interaction-complex}
# Fit complex regression
gat <- nlj::lm.gat(qsec ~ hp * mpg * wt, mtcars,
                   iterations = 6, penalty = 1e-9)

summary(gat$fit)
```
