---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
set.seed(420)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 300,
  fig.path = "man/figures/",
  fig.width = 9,
  fig.height = 6,
  out.width = "70%"
)
```

# NLJ

The `nlj` package, named in honor of British statistician Norman Lloyd Johnson,
provides specialized functions for working with the Johnson-SU (unbounded
system) distribution. In addition to functions for Johnson-SU distribution
computations, the package includes automatic normalization methods and a
generalized transformation-based regression model leveraging the inverse
hyperbolic sine transformation.

## Installation

To install the development version of `nlj`, use the following commands:

``` r
# install.packages("devtools")
devtools::install_github("shanedrabing/nlj")
```

## Overview of Functions and Examples

The `nlj` package provides functions for working with Johnson-SU distribution
metrics, normalization utilities, and advanced regression models. Below are
examples that illustrate the primary functions available in the package.

### Johnson-SU Distribution Functions

The Johnson-SU distribution functions (`djohnson`, `pjohnson`, `qjohnson`, and
`rjohnson`) allow for density, cumulative distribution, quantile, and random
variate generation. Each function has built-in parameterization for the
Johnson-SU distribution's unique flexibility in skewness and kurtosis
adjustments.

#### Density

This function calculates the density of the Johnson-SU distribution for a given
range of values.

```{r djohnson}
x <- seq(-pi, pi, length.out = 300)
d <- nlj::djohnson(x, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
plot(x, d, type = "l",
     main = "Johnson-SU Density",
     xlab = "X", ylab = "Density")
```

#### Cumulative Distribution

Use `pjohnson` to compute cumulative probabilities, allowing the mapping of
quantiles within the distribution.

```{r pjohnson}
q <- seq(-pi, pi, length.out = 300)
p <- nlj::pjohnson(q, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
plot(q, p, type = "l",
     main = "Johnson-SU Cumulative Distribution",
     xlab = "Quantile", ylab = "Probability")
```

#### Quantile Function

This function provides the quantile values for given cumulative probabilities,
useful for probabilistic analysis within the Johnson-SU framework.

```{r qjohnson}
p <- seq(0, 1, length.out = 300)
q <- nlj::qjohnson(p, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
plot(p, q, type = "l",
     main = "Johnson-SU Quantile",
     xlab = "Probability", ylab = "Quantile")
```

#### Random Deviate Generation

Generate random deviates from the Johnson-SU distribution with `rjohnson`,
suitable for simulation or resampling needs.

```{r rjohnson}
nlj::rjohnson(5, gamma = -1.7, delta = 2.1, xi = -1.45, lambda = 1.45)
```

### Automatic Normalization Functions

The package includes `znorm` and `zjohnson` functions, offering automatic
normalization, parameter binding, and denormalization capabilities. These
normalization utilities are particularly helpful in preparing data for analysis
or transformation to standardized forms.

#### Johnson-SU Normalization

The `zjohnson` function applies Johnson-SU normalization, suitable for data
that may not conform to a Gaussian distribution. This function optimizes
Johnson-SU parameters to best match the input data's distribution, offering a
transformation that can handle skewed and kurtotic data structures.

```{r johnson-density}
# Example data: Wind speed
x <- sort(airquality$Wind)
d <- density(x)

# Apply Johnson-SU normalization
zj <- nlj::zjohnson(x)

# Extract the density estimates
xd <- d$x
yd <- d$y
yj <- zj$fd(xd)

# Plot the density comparison
plot(range(xd), range(c(yd, yj)), type = "n",
     main = "Wind Speed Density Estimation",
     xlab = "Wind Speed", ylab = "Density")
lines(xd, yd, col = "black")  # Original density
lines(xd, yj, col = "red")   # Johnson-SU normalized density
legend(min(xd), max(c(yd, yj)),
       c("Kernel", "Johnson-SU"),
       col = c("black", "red"),
       lwd = 1, bg = "white")
```

### Generalized Asinh Transformation Model (GATM)

The Generalized Asinh Transformation Model (`lm.gat`) provides a flexible
approach to regression, using the generalized inverse hyperbolic sine
transformation. This model can improve fit quality by adapting to non-linearity
in data, especially when working with data where simple linear models fall
short.

#### Automatic Relationship Optimization

This example demonstrates the `lm.gat` function's ability to fit a non-linear
model to data by leveraging the generalized asinh transformation. Here,
horsepower (`hp`) is used to predict miles per gallon (`mpg`), and both a
simple and GATM model are fitted to highlight the latter's ability to capture
more complex patterns.

```{r gatm}
# Example data: Horsepower and miles per gallon (MPG)
i <- order(mtcars$hp)
x <- mtcars$hp[i]
y <- mtcars$mpg[i]

# Fit simple regression
m <- lm(y ~ x)

# Fit non-linear model
gat <- nlj::lm.gat(y ~ x, iterations = 3)

# Plot
plot(x, y, pch = 16,
     main = "Horsepower (HP) vs Miles Per Gallon (MPG)",
     xlab = "HP", ylab = "MPG")
lines(x, m$fitted.values, col = "red")
lines(x, gat$z, col = "blue")
legend(min(x), max(y),
       c("Simple", "GATM"),
       col = c("red", "blue"),
       lwd = 1, bg = "white")
```

#### Complex Nonlinear Interaction Optimization

In this example, `lm.gat` is used to fit a model with multiple interacting
terms. This showcases the function's capacity to handle complex,
high-dimensional relationships where interactions between variables (e.g.,
`hp`, `mpg`, `wt`) impact the outcome (`qsec`). This process emphasizes
`lm.gat`'s flexibility for optimizing non-linear and interaction-heavy models.

To begin, let's examine the results of naive linear regression model.

```{r interaction-simple}
# Fit simple regression
m <- lm(qsec ~ hp * mpg * wt, mtcars)

summary(m)
```

In the next code block, we apply `lm.gat` to fit a more flexible, non-linear
model. Unlike the previous linear model, `lm.gat` uses a generalized asinh
transformation to better capture non-linear interactions in the data. This
approach is particularly effective for capturing complex relationships.

```{r interaction-complex}
# Fit complex regression
gat <- nlj::lm.gat(qsec ~ hp * mpg * wt, mtcars,
                   iterations = 6, penalty = 1e-9)

summary(gat$fit)
```
